Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-10-14 Radio_Propagation_Simulation.tex
edits to thesis from clarice and modify background to reflect bistatic

\section{Introduction}
%introduce the section and what will be discussed here
The simulation software is broken down into three main components, the radio wave propagation model, the scene model, and the ray tracing engine. The wave propagation model is a collection of equations that describe the physical phenomena that affect the propagation when represented by rays. The model used is based upon classical geometric optics to work with the ray tracing engine. The scene model is comprised of a three dimensional environment in order to define the physical rotor blade surface. The receiver and transmitter will also be defined in the same environment with simulated characteristics to match real life conditions. The ray tracing engine uses recursive execution described in \cite{Whitted1979}(Whitted, 1979), replacing cameras and lights for receivers and transmitters respectively in a three dimensional environment.

\section{Radio Propagation Model}
%define the math behind the physical laws that are simulated within the simulation
The propagation model defined in this project is based on the following physical phenomena, free-space path loss, reflection, and the doppler effect. They are all formally established models described in full in \cite{Bertoni1999}, \cite{Parsons2000}, and \cite{Willis2005} (Bertoni, 2000; Parsons, 2000; Willis, 1991)(OK). Although several effects are described within the model several were left out to simplify the computation due to the overhead of implementation in three dimensional space and because of the assumptions made about the scene model.

@@

%formula description
where $\theta_i$ is the angle of incidence with the object's surface normal, $\varepsilon$ is the dielectric constant for that surface, and $\theta_t$ is the angle of refraction as described by Snell's law \ref{eqn:snell}.

@@

The final reflected power is

@@

%reflection power loss description
where $P_i$ is the incident power before reflection losses.

\subsubsection{Diffuse Reflection}
A perfect diffuse reflection, also known as Lambertian reflection, will reflect rays in all directions in a uniform hemisphere (J. Lambert, "Photometria Sive de Mensura et Gradibus Luminus Colorum et Umbrae", 1760). The hemisphere is located at the point of the ray hit, producing rays away from the surface. To approximate a particular type of surface the rays power is attenuated by the surface's bi-directional reflectance distribution function (BRDF) \cite{Suffern2007}, \cite{Pharr2010}, \cite{Glassner1995} (Suffern, 2007; Pharr, 2010; Glassner, 1995)(OK). There are several ways of calculating the BRDF, but the simplest is the Phong BRDF which will be used in this ray tracing implementation \cite{Phong:1998:ICG:280811.280980} (Phong, 1998)(OK). This method for the BRDF is regularly used in the rendering of computer graphics to simulate the reflection of light off surfaces. Since is was not designed for longer wavelengths used in RF communications an assumption is made that small scale features of the surface, relevant to light waves,  produce a reflectance distribution similar for larger wavelengths.

@@

It is assumed that both the receiver and the transmitter are not moving, therefore the only doppler imparted on the signal will be from the rotation of the blades. The associated Doppler is calculated using

@@

where $f_0$ is the initial transmitted frequency, $lambda$ is the wavelength, $v_t$ is the velocity in the direction of the transmitter, and $v_r$ is the velocity in the direction of the reflection. This is similar to the Bistatic RADAR equation for doppler, but since the reflection is not guaranteed to be toward the receiver we calculate the doppler as it makes contact with the surface and as it is re-radiated according the blades direction of travel.

@@

\section{Scene Model}
% describe what the scene model is and how it will fit in to the ray tracer and the 3d setting
The scene model is the 3D environment in which the physical objects are placed. 
The model is comprised of three types of objects, the rotor blades, transmitter and receiver. The rotor blade is the most complex object due to its airfoil shape. The transmitter is a point source that projects the rays into the scene, and the receiver is described in a similar fashion with a surrounding boundary layer. There are no other objects within the scene, this is because the effect of the rotor on the received signal is to be analyzed for defining characteristics in the best case scenario.

\subsection{Rotor Blade}
%go over the shape of the blade
The rotor blade is the most complex object within the scene due to its airfoil shape. The airfoil shape is designed to produce lift for the helicopter, and designing one is outside the scope of this project. Fortunately there are databases that provide real helicopter airfoil designs in x,y coordinates. Theses designs form a 2D cross-section of the airfoil shape that will be extruded into the length of the rotor blade. Figure \ref{fig:airfoil} shows the airfoil cross-section that will be used to create the rotor surface.

@@

%go over how it is formed and the tessellation code
The 2D airfoil sections will form the ribs of the rotor blade but the surface of the blade will be made up of primitive surfaces to approximate the curved airfoil shape. The primitive shapes are triangles that span between the ribs and go around to cover the surface. This was done using a lightweight mesh generation method using the a priori knowledge of the blade shape. The designed algorithm, located in \ref{class_blade__surface_a160959f632ad73eff846aeea81aa84ae} $create\_surface()$ method, takes advantage of the blade being composed of rib sections with consistent shape. The surface normals are then adjusted to match the correct directionality. Figure \ref{fig:tessilation} shows the created surface for the rotor blade that will be used in the scene.

@@

\subsection{Transmitter}
%point source transmitter casting rays in specific direction
The transmitter is modeled as a point source, this means that all the radio waves are projected from this point into the scene. It is assumed that the transmitter is located on the ground within the x,y plane in relation ship to the rotor blades. The transmitter is configured with a center frequency and is assumed to produce a tone at that frequency. Therefor signal produced has no inherent modulation applied to it.
%direction creation
Being that the Transmitter is emitting radio waves into a 3D space it is assumed to be transmitting in an omnidirectional fashion, but because the rotor and the receiver are the only two other objects within the scene radio waves will only be propagated in their direction. This limitation allows for a higher concentration of rays to be casted, resulting in a higher resolution picture and eliminates computational overhead for the tracing engine. Figure \ref{fig:transmitter_direction} shows the rays, in blue, casted only in the vicinity of the rotor blade and receiver.

@@

\section{Ray Tracing}
%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the scene and propagation models. This method of raytracing is established for creating realistic images in \cite{Suffern2007}, \cite{Pharr2010} (Suffern, 2007; Pharr, 2010). The simulation is performed over successive iterations. Each iteration is a discrete element called a frame. Where each frame will be updated to represent the change in the position of the rotor blade.

\subsection{Computation}
%flow of the program
The ray tracing engine first sets up the environment creating the surfaces and placing the objects in the scene. Then the number of frames that will be computed. The number of frames is determined by the number of rotations that the rotor makes within the simulation, the speed of the rotor in RPM and the sampling frequency of the receiver defined before the beginning of the simulation. This is described by

@@

The casting technique reduces the computation overhead by limiting the search space. It also creates higher resolution result by focusing the limited number of rays in the direction of the other objects shown in figure \ref{fig:samp_method}.

@@

When a ray is casted into the scene it has an origin point, a direction vector, a frequency and an initial power level. The ray then travels through the scene till it encounters another object. If there is a collision with the receiver it is recorded. If the ray collides with a rotor blade a new ray is calculated based on the normal to the geometric surface. The new ray's power is updated as well as an updated frequency since the blade is moving. The power level is adjusted based on the BRDF of the specific material. Due to only the rotor and receiver in the scene, besides the transmitter, rays that are found to hit the rotor the reflection ray is calculated. Then the angle between the reflection and a ray between the point of reflection and the receiver is used to create a dedicated ray in that direction with a decayed power and adjusted frequency accordingly.

\subsection{Output}
%iq samples
The superposition of these receiver contributions during each frame creates the received signal data. The received signal data will then be visualized in spectrograms over the length of the simulated time. The simulations do not render in real time but creates a CSV file that can then be loaded into MATLAB for signal processing.

>>>>

\section{Introduction}
%introduce the section and what will be discussed here
The simulation software is broken down into three main components, the radio wave propagation model, the scene model, and the ray tracing engine. The wave propagation model is a collection of equations that describe the physical phenomena that affect the propagation when represented by rays. The model used is based upon classical geometric optics to work with the ray tracing engine. The scene model is comprised of a 3D environment in order to define the physical rotor blade surface. The receiver and transmitter will also be defined in the same environment with simulated characteristics to match real life conditions. The ray tracing engine uses recursive execution described in \cite{Whitted1979}, replacing cameras and lights for receivers and transmitters respectively in a three dimensional environment.

\section{Radio Propagation Model}
%define the math behind the physical laws that are simulated within the simulation
The propagation model defined in this project is based on the following physical phenomena; free-space path loss, reflection, and the doppler effect. They are all formally established models described in full in \cite{Bertoni1999}, \cite{Parsons2000}, and \cite{Willis2005}. Although several effects are described within the model several were left out to simplify the computation due to the overhead of implementation in 3D space and because of the assumptions made about the scene model.

@@

%formula description
where $\theta_i$ is the angle of incidence with the object's surface normal, $\varepsilon$ is the dielectric constant for that surface, and $\theta_t$ is the angle of refraction as described by Snell's law

@@

The final reflected power is then

@@

%reflection power loss description
where $P_i$ is the incident power before reflection loss.

\subsubsection{Diffuse Reflection}
A perfect diffuse reflection, also known as Lambertian reflection, will reflect rays in all directions in a uniform hemisphere \cite{Sizun2005}. The hemisphere is centered at the point the ray hits, producing rays away from the surface. To approximate a particular type of surface the rays power is attenuated by the surface's bi-directional reflectance distribution function (BRDF) \cite{Suffern2007}, \cite{Pharr2010}, \cite{Glassner1995}. There are several ways of calculating the BRDF, but the simplest is the Phong BRDF which will be used in this ray tracing implementation \cite{Phong:1998:ICG:280811.280980}. This method for the BRDF is regularly used in the rendering of computer graphics to simulate the reflection of light off surfaces and combines the diffuse and specular components. Since this was not designed for the longer wavelengths used in RF communications, an assumption is made about surface features. Where features relevant to light waves,  produce a reflectance distribution similar for this application.

@@

It is assumed that both the receiver and the transmitter are not moving therefore, the only Doppler imparted on the signal will be from the rotation of the blades. The associated Doppler is calculated using

@@

where $f_0$ is the initial transmitted frequency, $\lambda$ is the wavelength, $v_t$ is the velocity in the direction of the transmitter, and $v_r$ is the velocity in the direction of the reflection. This is similar to the Bistatic radar equation for Doppler, but since the reflection is not guaranteed to be toward the receiver we calculate the doppler as it makes contact with the surface and as it is re-radiated according the blades direction of travel.

@@

\section{Scene Model}
% describe what the scene model is and how it will fit in to the ray tracer and the 3d setting
The scene model is the 3D environment in which the physical objects are placed. 
The model is comprised of three types of objects; the rotor blades, transmitter and receiver. The rotor blade is the most complex object due to its airfoil shape. The transmitter is a point source that projects the rays into the scene, and the receiver is described in a similar fashion with a surrounding boundary layer. There are no other objects within the scene. This is because the effect of the rotor on the received signal is analyzed for defining characteristics without outside variables.

\subsection{Rotor Blade}
%go over the shape of the blade
The rotor blade is the most complex object within the scene, due to its airfoil shape. The airfoil shape is designed to produce lift for the helicopter, and designing one is outside the scope of this project. Fortunately, there are databases that provide real helicopter airfoil designs in x,y coordinates. This data forms a 2D cross-section of the airfoil shape that will be extruded into the length of the rotor blade. Figure \ref{fig:airfoil} shows the airfoil cross-section that will be used to create the rotor surface.

@@

%go over how it is formed and the tessellation code
The 2D airfoil sections will form the ribs of the rotor blade but the surface of the blade will be made up of primitive surfaces to approximate the curved airfoil shape. The primitive shapes are triangles that span between the ribs and go around to cover the surface. This was done using a lightweight mesh generation method using a priori knowledge of the blade's shape. The designed algorithm, located in \ref{class_blade__surface_a160959f632ad73eff846aeea81aa84ae} $create\_surface()$ method, takes advantage of the object's uniform structure in 2D. The surface normals are then adjusted to match the correct directionality. Figure \ref{fig:tessilation} shows the created surface for the rotor blade that will be used in the scene.

@@

\subsection{Transmitter}
%point source transmitter casting rays in specific direction
The transmitter is modeled as a point source, this means that all the radio waves are projected from this point into the scene. It is assumed that the transmitter is located on the ground within the x,y plane in relationship to the rotor blades. The transmitter is configured with a center frequency and is assumed to produce a tone at that frequency. Therefor, the signal produced has no inherent modulation applied to it.
%direction creation
Being that the transmitter is emitting radio waves into a 3D space, it is assumed to be transmitting in an omnidirectional fashion. But because the rotor and the receiver are the only two other objects within the scene, radio waves will only be propagated in their direction. This limitation allows for a higher concentration of rays to be casted, resulting in a higher resolution picture that eliminates computational overhead for the tracing engine. Figure \ref{fig:transmitter_direction} shows the rays, in blue, casted only in the vicinity of the rotor blade and receiver.

@@

\section{Ray Tracing}
%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the scene and propagation models. This method of raytracing is established for creating realistic images in \cite{Suffern2007}, \cite{Pharr2010}. The simulation is performed over successive iterations. Each iteration is a discrete element called a frame and each frame will be updated to represent the change in the position of the rotor blade.

\subsection{Computation}
%flow of the program
The ray tracing engine first sets up the environment, creating the surfaces and placing the objects in the scene. Then the number of frames that will be computed. The number of frames is determined by the number of rotations that the rotor makes within the simulation, the speed of the rotor in RPM and the sampling frequency of the receiver defined before the beginning of the simulation. This is described by

@@

The casting technique reduces the computation overhead by limiting the search space. It also creates a higher resolution result by focusing the limited number of rays in the direction of the other objects shown in figure \ref{fig:samp_method}.

@@

When a ray is casted into the scene it has an origin point, a direction vector, a frequency, and an initial power level. The ray then travels through the scene till it encounters another object. If there is a collision with the receiver it is recorded. If the ray collides with a rotor blade a new ray is calculated based on the normal of the geometric surface. The new ray's power is then updated as well as an updated frequency since the blade is moving. The power level is adjusted based on the BRDF of the specific material. Due to only the rotor and receiver being in the scene, rays that are found to hit the rotor are reflected, then a dedicated ray in the direction of the receiver is also calculated with a decayed power and adjusted frequency respective to the reflection.

\subsection{Output}
%iq samples
The superposition of these receiver contributions during each frame creates the received signal data. The received signal data will then be visualized in spectrograms over the length of the simulated time. The simulation does not render in real time but instead creates a Comma Separated Value (CSV) file of In-phase and Quadrature (IQ) data that can then be loaded into MATLAB for signal processing.

### 2016-10-14 Ray_Tracing_Software.tex
added bib and modified for bistatic doppler calculation.
edits to thesis from clarice and modify background to reflect bistatic

The ray tracing software developed during the course of the thesis is written in C++ in an XCode environment, but a generic implementation is provided to run on other operating systems. This code is loosely based on the work of Eric Degroot's developed in (Eric's thesis) and implemented some modified base classes developed by him. It is a specific tool for analyzing the affects of rotor blade modulation on a signal tone, it does not provide generic 3D ray tracing in order to optimize the performance for this task. The software is provided without a GUI implementation and is run on different parameters by modifying the main.c file. This is due to the amount of parameters involved but could be easily extended to provide command line arguments in future iterations.

@@

the other key parameters that define the simulation are shown in table \ref{tab:key_values}

@@

The output file contains more than just the signal data the first few data entries are data pertaining to the simulation parameters. it is important to separate this data from the signal data before processing the signal. The template for this data is shown in table 
\ref{tab:file_values}.

>>>>

The ray tracing software developed during the course of the thesis is written in C++ in an XCode environment, but a generic implementation is provided to run on other operating systems. This code is loosely based on the work of Eric Degroot's developed for \cite{Groot2015} and implemented some modified base classes developed by him. It is a specific tool for analyzing the affects of rotor blade modulation on a signal tone, it does not provide generic 3D ray tracing in order to optimize the performance for this task. The software is provided without a GUI implementation and is run on different parameters by modifying the main.c file. This is due to the amount of parameters involved but could be easily extended to provide command line arguments in future iterations.

@@

the other key parameters that define the simulation are shown in Table \ref{tab:key_values}

@@

The output file contains more than just the signal data the first few data entries are data pertaining to the simulation parameters. it is important to separate this data from the signal data before processing the signal. The template for this data is shown in Table 
\ref{tab:file_values}.

### 2016-10-14 Results_new.tex
edits to thesis from clarice and modify background to reflect bistatic

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which at a set angle a doppler envelope is calculated and subsequent maximum and minimum values are collected. After a complete rotation the doppler maximum and minimum values produce a sinusoidal pattern as shown in figure \ref{fig:tx_azimuth_angle_200} without the assumption of knowing the starting azimuth angle. With this data we need to find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum envelope data is superimposed onto the maximum envelope data as shown in figure \ref{fig:azimuth_estimation_max_vs_absMin}.

@@

From figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. By computing the difference \ref{eqn:difference}

@@

we produce an almost exact sinusoid centered around zero, shown in figure \ref{fig:azimuth_estimation_difference}.

@@

While figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself, therefore by computing the correct doppler to use from the $Max Upper Envelope$ and the $| Min Lower Envelope |$. This is done by evaluating the difference at each angle and if the difference is negative the $| Min Lower Envelope |$ is used and if positive the $Max Upper Envelope$.

@@

The correct Doppler is shown in figure \ref{fig:correct_fd} which displays the amount of doppler that is currently acting on the signal for that specific azimuth angle. The data gathered contains a global minimum at 90\textdegree \space and a second local minimum at 270\textdegree. The 90\textdegree \space can be found easily by taking the minimum of the correct Doppler signal

@@

But in order to find the local minimum at 270\textdegree \space we need to use a different method. To do this we can use the peak detection function within Matlab to locate both the 90\textdegree and 270\textdegree point. By taking the inverse of the correct Doppler and tuning the peak detection function with the maximum and minimum values to determine accurate prominence.

@@

From figure \ref{fig:peaks}  we see the two peaks denoted by the red x's, one at 90\textdegree \space and the other at  270\textdegree. Now we have a minimum value that should denote the 90\textdegree \space location and the two peak locations that denote the 90\textdegree \space 270\textdegree locations. Because it is possible that the minimum could find the wrong peak for the 90\textdegree \space location we can compare the peak locations on the average power data, that was used to find the successive average during data collection, to determine which location is actually 90\textdegree.

@@

Figure \ref{fig:ave_power}  shows the average power at each angle sample with the angles 90\textdegree \space and 270\textdegree denoted. From the plot we can see that the average power at 90\textdegree \space is much smaller than the the power at 270\textdegree, so this inequality will be used to resolve ambiguity in the azimuth angle.

@@

From table \ref{tab:az_error_and_percent} we can see that there is no error in the estimation of the 90\textdegree \space point over a variety of Altitude and elevation angles that were tested. But because of the algorithm could potentially pick the 270\textdegree we will analyze the difference in peak heights and the power difference to evaluate the estimation further.

@@

From table \ref{tab:peaks_and_power} we can see that at the larger elevation angles the peak percent difference are significant to determine which is correctly the 90\textdegree value, but from the Peak frequency difference at the large elevation angles the difference is in the tens of Hz so in practical applications it is possible to confuse the estimation by just taking the minimum frequency so other values need to play a role such as the power.

@@

where $R_{Estimate}$ is the estimated reflection radius at a particular azimuth, $BladeLength$ is the length of a single blade or radius of the rotor, and $f_{DopplerCorrect}$ is the correct doppler frequency for each azimuth angle $\theta_{Az}$.

@@

where $f_{Doppler_{R Estimate Max}}$ is the maximum amount of doppler that could happen at that location if the signal was reflected parallel to the direction of motion. $f_c$ is the center frequency of the signal, $c$ is the speed of light and $v_r$ is the radial velocity based of the $RPM$ defined by

@@

Then to find the elevation angle we look back at equation (-) where we then solve for the elevation angle 
%Elevation angle
\begin{equation}
	\theta_{El} = acos\left(\frac{f_{DopplerCorrect(\theta_{Az})}}{f_{Doppler_{R Estimate Max}}}\right)
	\label{eqn:elevation_angle}
\end{equation}

@@

Figure \ref{fig:elevation_comparason_azimuth_range} shows the estimation of the elevation angle $\theta_{El}$ with respect to the azimuth angle, what it shows is that the elevation angle remains constant with respect to azimuth angle. which means that the estimation in the reflection radius is correctly compensating for the azimuth angle. The percent error off the actual elevation angle is shown in figure \ref{fig:percent_error_elevation_azimuth_range} which is very small and consistent over the azimuth range.

@@

From table \ref{tab:elevation_error_percent} elevation estimation error is extremely small for the given actual elevation angles. But there is a trend that is evident in the data that as the elevation angle decreases the amount of error is increasing.

@@

This error is caused by a limitation on the estimation is that the doppler shift is limited by the maximum amount doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is calculated by

@@

then using that to find the maximum available doppler through

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree \space. The maximum and minimum envelope data is shown in figure \ref{fig:envelope_180deg}

@@

Using the estimation method described above, figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till about 18\textdegree \space or 20\textdegree \space where it becomes noisy. This is due to the inability of the envelope estimation to resolve the correct maximum doppler for that range. By going back to figure (anglerangerelation) we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

If we now consider a pitched blade the above method for estimating Azimuth and elevation becomes inaccurate due to the increased doppler between 90\textdegree and 270\textdegree and the decreased doppler on the other side. To correct this error we need to correct both the difference calculation and the correct doppler data accordingly. The uncorrected data is shown in figure \ref{fig:diff_pitch_noCorrection} and \ref{fig:dop_pitch_noCorrection} with percent azimuth error of 20\% and percent elevation error of 35\%.

@@

Comparing the envelope difference of no pitch and pitched blade, in figures \ref{fig:azimuth_estimation_difference} and \ref{fig:diff_pitch_noCorrection} respectively, we can see that the difference calculation of the pitch is a shifted up version of the one with no pitch. Therefore we can first correct this by re-centering the pitch difference according to

\begin{equation}
	A_{shift} = \frac{Max(Max Uppe rEnvelope) - |Min(Min Lower Envelope)|}{2}
	\label{eqn:pitch_shift}
\end{equation}

@@

%fix correct doppler for pitch
\begin{equation}
	f_{DopplerCorrect} = f_{DopplerCorrect} + A_{shift}cos\left(\theta_{Az}\right)
	\label{eqn:fd_max}
\end{equation}

Since $\theta_{Az}$ is unknown during the calculation of this equation the correct starting phase can be determined by matching it to the $EnvelopeDifference$ with the a priori information that the cosign should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to unpitched results.

>>>>

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which a Doppler envelope and subsequent maximum and minimum values are calculated. After a complete rotation the Doppler maximum and minimum values produce a sinusoidal pattern as shown in Figure \ref{fig:tx_azimuth_angle_200} without the assumption of knowing the starting azimuth angle. With this data we need to find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum envelope data is superimposed onto the maximum envelope data as shown in Figure \ref{fig:azimuth_estimation_max_vs_absMin}.

@@

From Figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. By computing the difference

@@

where $Max Upper Envelope$ is an array of the maximum value of each upper envelope calculation, and $Min Lower Envelope$ is an array of the minimum value of each lower envelope calculation. $Envelope Difference$ is an almost exact sinusoid centered around zero, shown in Figure \ref{fig:azimuth_estimation_difference}.

@@

While Figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself, therefore by computing the correct Doppler using the $Max Upper Envelope$ and the $| Min Lower Envelope |$. This is done by evaluating the difference at each angle and if the difference is negative the $| Min Lower Envelope |$ is used and if positive the $Max Upper Envelope$.

@@

The correct Doppler is shown in Figure \ref{fig:correct_fd} which displays the amount of Doppler that is currently acting on the signal for that specific azimuth angle. The data gathered contains a global minimum at 90\textdegree \space and a second local minimum at 270\textdegree. The 90\textdegree \space can be found easily by taking the minimum of the correct Doppler signal

@@

But in order to find the local minimum at 270\textdegree \space we need to use a different method. To do this we can use the peak detection function within MATLAB to locate both the 90\textdegree \space and 270\textdegree \space point. By taking the inverse of the correct Doppler and tuning the peak detection function with the maximum and minimum values to determine accurate prominence.

@@

From Figure \ref{fig:peaks}  we see the two peaks denoted by the red x's, one at 90\textdegree \space and the other at  270\textdegree. Now we have a minimum value that should denote the 90\textdegree \space location and the two peak locations that denote the 90\textdegree \space and 270\textdegree \space locations. Because it is possible that the minimum could find the wrong peak for the 90\textdegree \space location we can compare the peak locations on the average power data, that was used to find the successive average during data collection, to determine which location is actually 90\textdegree.

@@

Figure \ref{fig:ave_power}  shows the average power at each angle sample with the angles 90\textdegree \space and 270\textdegree \space denoted as red x's. From the plot we can see that the average power at 90\textdegree \space is much smaller than the the power at 270\textdegree, so this inequality will be used to resolve ambiguity in the azimuth angle.

@@

From Table \ref{tab:az_error_and_percent} we can see that there is no error in the estimation of the 90\textdegree \space point over a variety of Altitude and elevation angles that were tested. But because of the algorithm could potentially pick the 270\textdegree \space we will analyze the difference in peak heights and the power difference to evaluate the estimation further.

@@

From Table \ref{tab:peaks_and_power} we can see that at the larger elevation angles the peak percent difference are significant to determine which is correctly the 90\textdegree \space value, but from the Peak frequency difference at the large elevation angles the difference is in the tens of Hz so in practical applications it is possible to confuse the estimation by just taking the minimum frequency so other values need to play a role such as the power.

@@

where $R_{Estimate}$ is the estimated reflection radius at a particular azimuth, $BladeLength$ is the length of a single blade or radius of the rotor, and $f_{DopplerCorrect}$ is the correct Doppler frequency for each azimuth angle $\theta_{Az}$.

@@

where $f_{Doppler_{R Estimate Max}}$ is the maximum amount of Doppler that could happen at that location if the signal was reflected parallel to the direction of motion. $f_c$ is the center frequency of the signal, $c$ is the speed of light and $v_r$ is the radial velocity based of the $RPM$ defined by

@@

Then to find the elevation angle we look back at equation \ref{eq:theory_180_upper} swapping $MaxUpperEnvelope$ with $f_{DopplerCorrect(\theta_{Az})}$ because this is now over the azimuth range and swap $f_{DopplerMax}$ for $f_{Doppler_{R Estimate Max}}$ because we need the maximum doppler for the current reflection estimate. we then solve for the elevation angle

%Elevation angle
\begin{equation}
	\theta_{El} = acos\left(\frac{f_{DopplerCorrect(\theta_{Az})}}{f_{Doppler_{R Estimate Max}}}\right)
	\label{eqn:elevation_angle}
\end{equation}

@@

Figure \ref{fig:elevation_comparason_azimuth_range} shows the estimation of the elevation angle $\theta_{El}$ with respect to the azimuth angle, what it shows is that the elevation angle remains constant with respect to azimuth angle. which means that the estimation in the reflection radius is correctly compensating for the azimuth angle. The percent error off the actual elevation angle is shown in Figure \ref{fig:percent_error_elevation_azimuth_range} which is very small and consistent over the azimuth range.

@@

From Table \ref{tab:elevation_error_percent} elevation estimation error is extremely small for the given actual elevation angles. But there is a trend that is evident in the data that as the elevation angle decreases the amount of error is increasing.

@@

This error is caused by a limitation on the estimation is that the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is calculated by

@@

then using that to find the maximum available Doppler through

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree \space. The maximum and minimum envelope data is shown in Figure \ref{fig:envelope_180deg}

@@

Using the estimation method described above, Figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till about 18\textdegree \space or 20\textdegree \space where it becomes noisy. This is due to the inability of the envelope estimation to resolve the correct maximum Doppler for that range. By going back to Figure \ref{fig:tx_range_elevation_rel} we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

If we now consider a pitched blade the above method for estimating Azimuth and elevation becomes inaccurate due to the increased Doppler between 90\textdegree \space and 270\textdegree \space and the decreased Doppler on the other side. To correct this error we need to correct both the difference calculation and the correct Doppler data accordingly. The uncorrected data is shown in Figure \ref{fig:diff_pitch_noCorrection} and \ref{fig:dop_pitch_noCorrection} with percent azimuth error of 20\% and percent elevation error of 35\%.

@@

Comparing the envelope difference of no pitch and pitched blade, in Figures \ref{fig:azimuth_estimation_difference} and \ref{fig:diff_pitch_noCorrection} respectively, we can see that the difference calculation of the pitch is a shifted up version of the one with no pitch. Therefore we can first correct this by re-centering the pitch difference according to

\begin{equation}
	A_{shift} = \frac{Max(Max Upper Envelope) - |Min(Min Lower Envelope)|}{2}
	\label{eqn:pitch_shift}
\end{equation}

@@

%fix correct doppler for pitch
\begin{equation}
	f_{DopplerCorrect} = f_{DopplerCorrect} + A_{shift}cos\left(\theta_{Az}\right)
	\label{eqn:fd_pitch_correct}
\end{equation}

Since $\theta_{Az}$ is unknown during the calculation of this equation the correct starting phase can be determined by matching it to the $EnvelopeDifference$ with a priori information that the cosign should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to Figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to the unpitched results.
